\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-26 body.tex
Adding related-work section.

@@

\section{Metatheory in (pseudo-)Agda}
We make three agda source code files (publicly and anonymously) openly avaliable that include the corresponding QIIT syntax definition and models.

For syntax, please check \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{here}.

For consistency model, please check \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{here}.

For canonicity model, please check \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{here}.

We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins.

>>>>

\section{Related Work}\label{sec:related-work}
\input{related-work.tex}

@@

\section{Metatheory in (pseudo-)Agda}
We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.

For syntax, please check \\ \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}. This is done by mainly following the formulation \citet{altkap2016}. This is also very close to \citet{coquand2018canonicity} and \citet{sterling2019algebraic} but we formulate our syntax in a type-theoretic framework.

For syntactical translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}. This is the first QIIT-model we give. A QIIT-model can be consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case). Here we use (a subpart of) QIIT data itself as the destination of the mapping.

For consistency model, please check \\ \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}. This is following the standard model from \citet{altkap2016,kaposi2017type,kaposi2019gluing}.

For canonicity model, please check \\ \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}. This is mainly following \citet{coquand2018canonicity,sterling2019algebraic}, but we carried out our argument in a type-theortic framework.  We call it canonicity model to emphasize that it is a model used to prove canonicity, it is also called reducibility argument, as an instance of proof-relevant logical relation. This argument can be considered as a verbose/more concrete version of \citet{kaposi2019gluing}.

We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.

### 2022-06-26 metatheory.tex
Adding syntactical translation.

Though phrased in Latex, we are following~\citet{altkap2016} to use as
our meta-logic a type theoretical framework (like Agda), 
together with \textit{Quotient Inductive Inductive Type} (QIIT). Each
judgement will actually be represented by a QIIT type, which are
mutually recursively defined together. Quotient is used to represent
judgemental(definitional) equality, thanks to which we can have a
concise representation. This way, for all four kinds of judgement we are
using, there are always equality among these data. Doing so, we can omit
some obvious equality judgement. Most of the reasoning can be formulated
using the \textit{algebra} of QIIT (mapping out function from this
quotient data-type).

@@

>>>>

Though phrased in Latex, we are following~\citet{altkap2016} to use as
our meta-logic a type theoretical framework (like Agda), 
together with \textit{Quotient Inductive Inductive Type} (QIIT). Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}. We base our formulation of the syntax in this type theory instead of a more conventional and informal representation, becase (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out.

In our syntax, each judgement will actually be represented by a QIIT type, which are
mutually recursively defined together. Quotient is used to represent
judgemental(definitional) equality, thanks to which we can have a
concise representation. This way, for all four kinds of judgement we are
using, there are always equality among these data. Doing so, we can omit
some obvious equality judgement. Most of the reasoning can be formulated
using the \textit{algebra} of QIIT (mapping out function from this
quotient data-type).

@@

\subsection{Implementation Formalized as Syntactical Translation}

\input{syn-translate.tex}

### 2022-06-26 syn-translate.tex
Adding "accessing family components outside that family" into lang-design section Slight change on syntactic translation section.

The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory. The described compilation strategy can be considered as a map from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}. This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.

@@

We omit all the equational rules ($\beta,\eta$ and substitution) here. However, for coherence, when we mutual recursively define $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and $\denotesT{\cC}$ above, we actually have to prove the substitution laws $(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau] \equiv \Sigr{\Gamma}{n}$ together. The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.

@@

>>>>

The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory. The described compilation strategy can be considered as a \textit{syntactical translation}, a function mapping from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}.\footnote{Here, syntactical QIIT-model can be roughly understood as just AST, with equational rules acting like quotient. So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.

@@

We omit all the equational rules ($\beta,\eta$ and substitution) here. However, for coherence, when we mutual recursively define $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and $\denotesT{\cC}$ above, we actually have to prove the substitution laws $(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau] \equiv \Sigr{\Gamma}{n}$ together. The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form. We encourage the reader to read that (pseudo-)Agda-style proof because it is more friendly to type theorists, easy to do type checking and validate the proof and it is clearer on the mathematical object each translation uses.

@@

% the link for syntactical translation
% https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing
